---
title: "Pstat 131 Homework 5"
author: "Yu Tian"
date: "Spring 2022-05-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      results = 'markup',
                      fig.width = 4,
                      fig.height = 3,
                      fig.align = 'center',
                      message = F,
                      warning = F)

# packages
library(tidyverse)
library(tidymodels)
library(ISLR)
library(ISLR2)
library(yardstick)
library(readr)
library(discrim)
library(poissonreg)
library(corrr)
library(glmnet)
library(dplyr)
tidymodels_prefer()
```

## Elastic Net Tuning

#### View Pokemon Date
```{r}
# Read the Pokemon data set into R using read_csv()
Pokemon <- read_csv(file = "Pokemon.csv") 
Pokemon %>% head()

```

## Exercise 1
Install and load the janitor package. Use its clean_names() function on the Pokémon data, and save the results to work with for the rest of the assignment. What happened to the data? Why do you think clean_names() is useful?

#### Answer
```{r}
library(janitor)

pokemon <- Pokemon %>%
  clean_names()

head(pokemon)
```
Compared with the two tables above, we can find the cleam_names() function make the variables name changed with lower-case letter. Besides, the space and dots which separate the words becomes the underscore "_". Since this function standardizes all the variable names into the same clear format, it will be more convenient and easier for the latter understanding and data processing.


## Exercise 2
Using the entire data set, create a bar chart of the outcome variable, type_1.

How many classes of the outcome are there? Are there any Pokémon types with very few Pokémon? If so, which ones?

For this assignment, we’ll handle the rarer classes by simply filtering them out. Filter the entire data set to contain only Pokémon whose type_1 is Bug, Fire, Grass, Normal, Water, or Psychic.

After filtering, convert type_1 and legendary to factors.

#### Answer
```{r}
# Using the entire data set, create a bar chart of the outcome variable, type_1.
count = table(pokemon$type_1)
count
barplot(count, main="Pokemon Type 1", 
        xlaby="the type_1 of pokemon", ylab="the number of pokemon", 
        width = 1, cex.names = 0.3)
```



```{r}
# How many classes of the outcome are there? Are there any Pokémon types with very few Pokémon? If so, which ones?
pokemon %>%
  group_by(type_1) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

From the table above, we can find that there are 18 classes of the outcome. Flying is the Pokémon types with very few Pokémon whose count is only 4.

```{r}
# Filter the entire data set to contain only Pokémon whose type_1 is Bug, Fire, Grass, Normal, Water, or Psychic. 
pokemon_filter <- pokemon %>% filter(type_1 == "Bug" |
                                       type_1 == "Fire" |
                                       type_1 == "Grass" |
                                       type_1 == "Normal" |
                                       type_1 == "Water" |
                                       type_1 == "Psychic")
pokemon_filter %>%
  group_by(type_1) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

```{r}
# Convert type_1 and legendary to factors.
pokemon_filter_factor <- pokemon_filter %>%
  mutate(type_1 = factor(type_1)) %>%
  mutate(legendary = factor(legendary)) %>%
  mutate(generation = factor(generation))
head(pokemon_filter_factor)
```


## Exercise 3
Perform an initial split of the data. Stratify by the outcome variable. You can choose a proportion to use. Verify that your training and test sets have the desired number of observations.

Next, use v-fold cross-validation on the training set. Use 5 folds. Stratify the folds by type_1 as well. Hint: Look for a strata argument. Why might stratifying the folds be useful?

#### Answer
```{r}
set.seed(0623)
pokemon_split <- initial_split(pokemon_filter_factor, prop = 0.7, strata = type_1)
pokemon_train <- training(pokemon_split)
pokemon_test <- testing(pokemon_split)
dim(pokemon_filter_factor)
dim(pokemon_train)
dim(pokemon_test)
```

```{r}
# Verify the training and testing data sets have the appropriate number of observations
# the number of observations for all data
a <- nrow(pokemon_filter_factor)
a
# the number of observations for training data
b <- nrow(pokemon_train)
b
# the number of observations for test data
c <- nrow(pokemon_test)
c
# the percentage of observations for training data
b/a
# the percentage of observations for test data
c/a
```
The probability of training data observations is 0.6943231, which is almost equal to prob=0.70, so the training and testing data sets have the desired number of observations.

```{r}
# use v-fold cross-validation on the training set. Use 5 folds. Stratify the folds by type_1 as well.  Why might stratifying the folds be useful?
pokemon_folds <- vfold_cv(pokemon_train, v = 5, strata = type_1)
pokemon_folds 
```
We are trying to use v-fold cross-validation and divide the testing data into 5 groups of roughly equal size to prepare for the later fitting and prediction process. 

v-fold cross-validation is one kind of resampling method. For each model, this method will randomly divide the observation data into v groups of roughly equal sizes, which are folds. This method will hold out the 1st fold as the validation set to be evaluated. Then the remaining k-1 folds will be analyzed to fit the model. The final estimate of model will get by the average of k results.

Thus, it is useful to make sure the distribution of types in each fold is balanced with the entire data set.




## Exercise 4
Set up a recipe to predict type_1 with legendary, generation, sp_atk, attack, speed, defense, hp, and sp_def.

Dummy-code legendary and generation;

Center and scale all predictors.

#### Answer
```{r}
# Set up a recipe 
pokemon_recipe <- recipe(type_1 ~ legendary + generation + sp_atk + attack + 
                           speed + defense + hp + sp_def, data = pokemon_train) %>%
  step_dummy(legendary) %>%
  step_dummy(generation) %>%
  step_center(all_predictors()) %>% 
  step_scale(all_predictors())

pokemon_recipe
```




## Exercise 5
We’ll be fitting and tuning an elastic net, tuning penalty and mixture (use multinom_reg with the glmnet engine).

Set up this model and workflow. Create a regular grid for penalty and mixture with 10 levels each; mixture should range from 0 to 1. For this assignment, we’ll let penalty range from -5 to 5 (it’s log-scaled).

How many total models will you be fitting when you fit these models to your folded data?

#### Answer
```{r}
#set up model
pokemon_model <- multinom_reg(penalty = tune(), mixture = tune()) %>% 
  set_engine("glmnet")

# set up workflow
pokemon_workflow <- workflow() %>% 
  add_model(pokemon_model) %>% 
  add_recipe(pokemon_recipe)

# Create a regular grid 
pokemon_grid <- grid_regular(penalty(range = c(-5, 5)), mixture(range = c(0, 1)), 
                             levels = c(10, 10))
```

500 models in total will be fitting to the data. There are 5 folds and 100 models I will fit to each fold, so total number is 5*100=500. 





## Exercise 6
Fit the models to your folded data using tune_grid().

Use autoplot() on the results. What do you notice? Do larger or smaller values of penalty and mixture produce better accuracy and ROC AUC?

#### Answer

```{r}
# Fit the models to your folded data using tune_grid().
pokemon_tune_res <- tune_grid(pokemon_workflow,
                               resamples = pokemon_folds,
                               grid = pokemon_grid)

# Use autoplot() on the results.
autoplot(pokemon_tune_res)
```
What do you notice? Do larger or smaller values of penalty and mixture produce better accuracy and ROC AUC?
From the graph above, we can find that with the value of penalty and mixture get larger, the value of accuracy and roc_auc become smaller. Thus, The smaller values of penalty and mixture produce better accuracy and ROC AUC. 




## Exercise 7
Use select_best() to choose the model that has the optimal roc_auc. Then use finalize_workflow(), fit(), and augment() to fit the model to the training set and evaluate its performance on the testing set.

#### Answer
```{r}
# Use select_best() to choose the model that has the optimal roc_auc.
optimal_auc <- select_best(pokemon_tune_res, metric = "roc_auc")

# use finalize_workflow(), fit(), and augment() to fit the model to the training set and evaluate its performance on the testing set.
pokemon_final <- finalize_workflow(pokemon_workflow, optimal_auc)

pokemon_final_fit <- fit(pokemon_final, data = pokemon_train)

augment(pokemon_final_fit, new_data = pokemon_test) %>% 
  accuracy(truth = type_1, estimate = .pred_class)
```

The accuracy of testing set is 0.3142857, which performs not well.


## Exercise 8
Calculate the overall ROC AUC on the testing set.

Then create plots of the different ROC curves, one per level of the outcome. Also make a heat map of the confusion matrix.

What do you notice? How did your model do? Which Pokemon types is the model best at predicting, and which is it worst at? Do you have any ideas why this might be?

#### Answer
```{r}
# Calculate the overall ROC AUC on the testing set.
roc_auc(augment(pokemon_final_fit, new_data = pokemon_test), type_1, .pred_Bug, .pred_Fire, 
                .pred_Grass, .pred_Normal, .pred_Psychic, .pred_Water) 
```

```{r}
#  create plots of the different ROC curves, one per level of the outcome.
augment(pokemon_final_fit, new_data = pokemon_test) %>% 
  roc_curve(type_1, .pred_Bug, .pred_Fire, .pred_Grass, .pred_Normal, 
            .pred_Psychic,.pred_Water) %>%
  autoplot()
```

```{r}
# make a heat map of the confusion matrix.
augment(pokemon_final_fit, new_data = pokemon_test) %>%
  conf_mat(truth = type_1, estimate = .pred_class)  %>% 
  autoplot(type = "heatmap")
```



From the graphs and calculation results above, we can find that the value of accuracy and roc_auc are not very high. Thus, the model performs not very well. Also, since the prediction accuracy of the six types is different, we can find that Normal Pokemon type is the model best at predicting, the Water Pokemon type is the second model best at predicting, the Grass Pokemon type is the model worst at predicting, the Fire Pokemon type is the second model worst at predicting. I think we get this result because the Normal and Water Pokemon Type has more observations in the data set, and the Grass and Fire Pokemon Type has less observations to predict in the data set.


















